---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pipaux

<!-- badges: start -->
[![R-CMD-check](https://github.com/PIP-Technical-Team/pipaux/workflows/R-CMD-check/badge.svg)](https://github.com/PIP-Technical-Team/pipaux/actions?workflow=R-CMD-check)
[![Codecov test coverage](https://codecov.io/gh/PIP-Technical-Team/pipaux/branch/master/graph/badge.svg)](https://app.codecov.io/gh/PIP-Technical-Team/pipaux?branch=master)
<!-- badges: end -->

`pipaux` manages the auxiliary data used in the PIP workflow. It allows two main actions: 
[1] It efficiently updates the auxiliary data in the Y drive, while making sure it is in sync with the raw data in Github, and that the right dependencies are processed. It also logs the updates, so that users can inspect timing, success, and errors information.
[2] It tracks changes across and within releases 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("PIP-Technical-Team/pipaux")
```

```{r setup, eval = FALSE}
# library(pipaux)
devtools::load_all()
```


Even though `pipaux` has many functions, one per each auxiliary data measure, most of its features can be executed through the following key functions:
 
* `update_aux_measures()`: to update auxiliary data measures, either one at a time or all together, in the right order of dependencies.
* `pipload::load_aux_data()`: to load the auxiliary data into memory. This function is a wrapper around the measure-specific loading functions (e.g., `aux_cpi("load")`), and it allows users to load one or more measures together.
* `compare_aux_releases()`: to compare auxiliary data across releases, and detect new rows, removed rows, or changed values.
* `compare_aux_vintages()`: to compare auxiliary data within the same release, and detect version changes.

## Mandatory Setup (Working Release)

Before using any update or comparison functions, you must initialize the working release:

```{r, eval = FALSE}
pipfun::setup_working_release(
  release  = "20260202", #for example
  identity = "TEST"
)
```

This populates the internal .pipaux environment with:

* Release date (e.g. `YYYYMMDD`)

* Identity (e.g., `TEST`, `PROD`)

* Y-drive data path

* Y-drive metadata path

* Stamp aliases for versioned artifacts

You can inspect the active release:
```{r, eval = FALSE}
get_from_auxenv("wrk_release")
get_from_auxenv("aux_data_path")
```

## Core concepts

Although pipaux contains one `aux_*` function per measure, most workflows rely on these main functions:

**1. Load auxiliary data (read-only)**
```{r, eval = FALSE}
pipload::load_aux_data(measure = "gdp")
```

Loads previously saved auxiliary data from the Y drive, in the release specific folder.
This does not trigger an update.

**2. Update one measure**
```{r, eval = FALSE}
aux_fun(measure = "gdp", owner = "YourGHUser")
```

This:

* Checks whether the GitHub release branch is up to date

* Detects raw data SHA changes

* Detects formatter function code changes

* Resolves and updates dependencies automatically

* Saves new versions only if needed

**3. Update multiple measures in dependency-aware order**
```{r, eval = FALSE}
update_aux_measures(
  measures = c("cpi", "metaregion"),
  owner    = "YourGHUser",
  log      = TRUE,  
  log_save = FALSE,        # Optional: save log to disk
  halt_on_dep_fail = FALSE # Continue updates even if a dependency fails
)
```

This:

* Processes measures in dependency order

* Avoids unnecessary work

* Optionally saves a structured update log

**4. Compare auxiliary data across releases**
```{r, eval = FALSE}
compare_aux_releases(old_release = "YYYYMMDD_ID")
```

These functions identify:

* Added rows

* Removed rows

* Changed values

### How Updates Decide Whether to Run

`pipaux` avoids unnecessary recomputation using multiple checks:

* GitHub release branch vs `DEV`

* Raw file SHA changes

* `aux_*` function code hash changes

* Dependency cascade detection

* Presence and integrity of Y-drive sidecar metadata

If none of these signals change, the measure is not republished.

Versioning is handled via the stamp framework:

```{r, eval = FALSE}
pipaux_set_versioning("content")    # default
pipaux_set_versioning("timestamp")  # force new version every run
pipaux_set_versioning("off")        # overwrite without versioning
```

**Update Workflow (Visual Overview)**

```mermaid
flowchart TD
    A["aux_fun() or update_aux_measures()"] --> B[Init release + top-level log]
    B --> C[Resolve repo / owner]
    C --> D[Process dependencies recursively]
    D --> E[get_fs_status()]
    E --> F{GitHub update needed?}
    F -->|Yes| G["update_gh = TRUE<br>update_y = TRUE"]
    F -->|No| H[Check Y-drive status]
    H --> I{SHA mismatch<br>or code hash mismatch?}
    I -->|Yes| J[update_y = TRUE]
    I -->|No| K[No update needed exit]
    G --> L[execute_update()]
    J --> L
    L --> M{update_gh?}
    M -->|Yes| N[Sync release branch from DEV]
    M -->|No| O[Skip GH sync]
    N --> P{update_y?}
    O --> P
    P -->|Yes| Q["Call aux_<measure>(action = 'update')"]
    P -->|No| R[Skip Y update]
    Q --> S[Save artifact + sidecar metadata]
    S --> T[Log success]
```



### Logging

Every update call generates structured log entries including:

* Measure name

* Timestamp

* Success / failure

* GitHub SHAs

* Error stack traces (if any)

If `log_save = TRUE`, call `pipfun::log_load()`

If not saved, the most recent in-memory log can be accessed via `aux_log_last()`.

## Typical Workflow

```{r, eval = FALSE}
# 1. Initialize release
pipfun::setup_working_release()

# 2. Update selected measures
update_aux_measures(
  measures = c("cpi", "gdp"),
  owner = "YourGHUser",
  log_save = TRUE
)

# 3. Load updated data
gdp <- pipload::load_aux_data("gdp")

# 4. Compare to previous release
compare_aux_releases(old_release = "YYYYMMDD_TEST")
```

## Scope

*pipaux* is intended for:

* PIP Technical Team workflows

* Controlled release environments (TEST / PROD)

* Server environments with Y-drive access

* Users with valid GitHub PAT credentials

It is not intended as a general-purpose data pipeline framework.

## Further Documentation

For deeper technical details (dependency graph logic, sidecar metadata structure, SHA semantics, logging internals, development diagnostics), see the project technical guide in `dev/_project_notes`

## Summary

`pipaux` ensures that auxiliary data in the PIP workflow are:

* Reproducible

* Versioned

* Dependency-aware

* Provenance-tracked

* Efficiently updated

* Fully auditable

It is the orchestration layer between GitHub auxiliary repositories and the PIP Y-drive release system.