---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pipaux

<!-- badges: start -->
[![R-CMD-check](https://github.com/PIP-Technical-Team/pipaux/workflows/R-CMD-check/badge.svg)](https://github.com/PIP-Technical-Team/pipaux/actions?workflow=R-CMD-check)
[![Codecov test coverage](https://codecov.io/gh/PIP-Technical-Team/pipaux/branch/master/graph/badge.svg)](https://app.codecov.io/gh/PIP-Technical-Team/pipaux?branch=master)
<!-- badges: end -->

`pipaux` manages the auxiliary data used in the PIP workflow. It basically
does two things. [1] It updates the auxiliary data and makes sure to keep 
it unchanged when it the raw data has not been modified. [2] it loads the 
data into memory.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("PIP-Technical-Team/pipaux")
```

```{r, include = FALSE}

last_item <- function(x, word = "and") {
  if (!(is.character(x))) {
    warning("`x` must be character. coercing to character")
    x <- as.character(x)
  }

  lx <- length(x)
  if (lx == 1) {
    y <- x
  }
  else if (lx == 2) {
    y <- paste(x[1], word, x[2])
  }
  else {
    y <- c(x[1:lx-1], paste(word, x[lx]))
    y <- paste(y, collapse = ", ")
  }
  return(y)
}


```

```{r setup}
library(pipaux)
```

```{r numb-func, include = FALSE}
lf <- lsf.str("package:pipaux", pattern = "^pip_")
  lf <- as.character(lf)
  num_functions <- length(lf)
```

Even though `pipaux` has more than `r num_functions` functions, most of its features 
can be executed by only using functions `pipaux::load_aux` and `pipaux::update_all_aux`, which wrappers of other measure-specific functions. 

## Loading data

```{r functions-av, include = FALSE}
  lf <- lsf.str("package:pipaux", pattern = "^pip_[a-z]{3}$")
  lf <- as.character(lf)
  lf <- gsub("pip_", "", lf)
```

Loading auxiliary data can be done using the function `pipaux::load_aux`, whose 
first argument is the name of the `measure` to be loaded. The measures available
are **`r last_item(lf)`**. 

So, loading the `cpi` database could be done by typing, 
`load_aux(measure = "cpi")` or `aux_cpi("load")`. 

Remember to first set up your working release, so that `load_aux()` will automatically load the file in the current release folder. For example, 

```{r}
pipfun::setup_working_release(release = "20250203")
```


```{r equal-load, echo=TRUE}
identical(load_aux(measure = "cpi"), aux_cpi("load"))
identical(load_aux(measure = "ppp"), aux_ppp("load"))
identical(load_aux(measure = "pop"), aux_pop("load"))
```

## Updating data

Updating the auxiliary data in PIP can be done either one measure at a time or all measures together. 
Data could be updated using the function `pipaux::aux_fun(measure = "...")` of any of the measure 
specific functions. Or all auxiliary data measures can be updated altogether with `update_all_aux`. Both of these functions also update any dependencies of the specified measure.

Alternatively, updating the specified measure (e.g., cpi) only can be done using again the `aux_cpi(action = "update")` function.

```{r equal-update, eval = FALSE}
# this, 
aux_fun(measure = "cpi")

# is equivalent to this, 
aux_cpi("update")

# And this
update_all_aux(measure = "cpi")

```

`pipaux` checks whether the hash of each database has changed and updates the data 
in case it has changed. If it has not changed, it notifies the user. The user can 
force the update by using `force = TRUE` as part of the arguments.
