---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pipaux

<!-- badges: start -->
[![R-CMD-check](https://github.com/PIP-Technical-Team/pipaux/workflows/R-CMD-check/badge.svg)](https://github.com/PIP-Technical-Team/pipaux/actions?workflow=R-CMD-check)
[![Codecov test coverage](https://codecov.io/gh/PIP-Technical-Team/pipaux/branch/master/graph/badge.svg)](https://app.codecov.io/gh/PIP-Technical-Team/pipaux?branch=master)
<!-- badges: end -->

`pipaux` manages the auxiliary data used in the PIP workflow. It allows two main actions: 
[1] It efficiently updates the auxiliary data in the Y drive, while making sure it is in sync with the raw data in Github, and that the right dependencies are processed. It also logs the updates, so that users can inspect timing, success, and errors information.
[2] It tracks changes across and within releases 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("PIP-Technical-Team/pipaux")
```

```{r setup}
library(pipaux)
```

```{r numb-func, include = FALSE}
lf <- lsf.str("package:pipaux", pattern = "^aux_")
  lf <- as.character(lf)
  num_functions <- length(lf)
```

Even though `pipaux` has more than `r num_functions` functions, one per each auxiliary data measure, most of its features can be executed through the following key functions: 
 
`update_aux_measures()`: to update auxiliary data measures, either one at a time or all together, in the right order of dependencies.
`pipload::load_aux_data()`: to load the auxiliary data into memory. This function is a wrapper around the measure-specific loading functions (e.g., `aux_cpi("load")`), and it allows users to load one or more measures together.
`compare_aux_releases()`: to compare auxiliary data across releases, and detect new rows, removed rows, or changed values.
`compare_aux_vintages()`: to compare auxiliary data within the same release, and detect version changes.

## Preliminary step 
Setting up the working release to ensure all info is in place for the updates and comparisons to run smoothly. This includes defining the release name and identity (e.g., TEST, PROD).

## Overview of key outputs

## Loading data

## Updating data

## Comparing data versions 