---
title: "Aux Data - Pipeline Overview"
author: "PIP Technical Team"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aux Data - Pipeline Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
library(pipaux)
```

# Overview

This vignette provides a technical overview of the auxiliary data (aux data)
pipeline, detailing how measures are updated, logged, and compared across
releases and versions. It is intended for the PIP Technical Team to run
updates safely and inspect changes over time.

## Objectives

- Update auxiliary data measures in dependency-aware order.
- Synchronize data between GitHub and Y-drive.
- Inspect logs for timing, success, and error information.
- Optionally simulate changes for testing purposes.
- Compare data across releases or within vintages.

---

# Key Functions

| Function                 | Purpose                                                                                        |
|--------------------------|------------------------------------------------------------------------------------------------|
| `update_aux_measures()`  | Updates all or selected auxiliary data measures in dependency order. Handles GitHub/Y-drive updates, logging, and dependency resolution. |
| `aux_fun()`              | Called internally by `update_aux_measures()`. Executes updates for a single measure and manages its dependencies. |
| `compare_aux_releases()` | Compares aux data files across two releases to detect added/removed rows or changed values.    |
| `compare_aux_vintages()` | Compares aux data files within the same release across vintage versions.                       |
| `simulate_changes()`     | Testing only. Simulates row drops and value modifications in an aux data file.                 |

> `check_status()` and `get_fs_status()` are internal functions used by
> `aux_fun()` to determine whether GitHub or Y-drive updates are needed.
> They are not intended for direct use in routine workflows.

---

# Key Outputs

- Updated GitHub release branches (`release_branch = <release>_<identity>`)
  for each measure.
- Synchronized auxiliary data files saved to the Y-drive
  (`getOption("pipaux.working_dir")`).
- Logs capturing:
    - Measure name
    - Step type (`START`, `UPDATE_GH`, `UPDATE_Y`, `END`, `ERROR_*`)
    - Timing and status
    - Optional error messages

---

# 1️⃣ Step 1 – Set up the Working Release

All pipeline operations are tied to a working release, which defines the
release name and identity (e.g., `TEST`). Initialize at the start of each
session:

```{r}
pipfun::setup_working_release(
  release  = "20260202",  # e.g. YYYYMMDD
  identity = "TEST"
)
```

---

# 2️⃣ Step 2 – Update Auxiliary Data Measures

Update all measures or a subset in dependency-aware order using
`update_aux_measures()`. Logs are automatically captured.

```{r}
# Update specific measures with logging enabled
update_aux_measures(
  measures         = c("cpi", "metaregion"),
  owner            = "RossanaTat",
  log              = TRUE,
  log_save         = FALSE,       # set TRUE to persist log to disk
  halt_on_dep_fail = FALSE        # continue even if a dependency fails
)

# Update ALL available measures (dependency order is resolved automatically)
update_aux_measures(
  owner    = "RossanaTat",
  log      = TRUE,
  log_save = TRUE
)
```

## Notes on Updates

- **Dependency handling:** Measures are updated in the correct order based
  on the shared dependency manifest. You do not need to manage ordering
  manually.
- **GitHub/Y-drive updates:** The `update_gh` and `update_y` flags are
  determined automatically by `get_fs_status()` — no need to set them
  directly. A GitHub update always implies a Y-drive update.
- **`measures = NULL`:** Passing `NULL` (the default) updates all available
  measures in dependency order.

## Inspecting the Log

```{r}
# Option 1: retrieve the last log from the in-memory logging environment
last_log <- aux_log_last()

# Option 2: filter for update events only (using pipfun directly)
pipfun::log_get("pipaux_update_log")[event == "update"]

# Option 3: load a previously saved log from disk (requires log_save = TRUE)
pipfun::log_load(
  id    = "pipaux_update_log",
  alias = get_from_auxenv("aux_meta_alias")
)
```

> `aux_log_last()` is an internal helper that returns the most recent log
> object from the `.piplogenv` environment. `aux_log_last_name()` returns
> its name as a character string.

---

# 3️⃣ Step 3 – Simulate Changes (Testing Only)

`simulate_changes()` lets you introduce hypothetical row drops and value
modifications into an aux data file. The modified file is saved to the
Y-drive and can then be used to test the comparison functions in Step 4.

```{r}
# Simulate changes for a single measure
simulate_changes(
  measure = "cpi",    # scalar — one measure at a time
  drop    = c(1, 2),  # row indices to drop
  indices = c(1, 3),  # row indices to modify values
  seed    = 123
)
```

> **Note:** `measure` takes a single character string, not a vector.
> Call `simulate_changes()` once per measure.

**Important:** Simulations are only for testing pipeline logic. During a
production workflow, skip this step entirely.

---

# 4️⃣ Step 4 – Compare Auxiliary Data

Once updates (or simulations) are complete, inspect changes across releases
or within vintages using the comparison functions.

## Compare Across Releases

`compare_aux_releases()` compares the current working release against a
prior release for one or more measures. If `old_release` is not specified,
it automatically identifies the most recent prior release with the same
identity.

```{r}
# Compare a single measure (old_release auto-detected)
changes_pop <- compare_aux_releases(
  measure = "pop"
)

# Compare multiple measures against a specific prior release
changes <- compare_aux_releases(
  measure     = c("cpi", "gdp"),
  old_release = "20260101_TEST",
  verbose     = TRUE
)
```

The result is a named list (one element per measure). Each element contains:

- `diff_values`: data table of value-level differences (`NULL` if none).
- `diff_rows`: data table of added/removed rows (`NULL` if none).
- `attr(result$<measure>, "key_cols")`: primary key columns used for
  comparison (stored as an attribute, not a list element).

```{r}
# Access key columns used for a specific measure
attr(changes$cpi, "key_cols")
```

## Compare Within the Same Release (Vintage Versions)

`compare_aux_vintages()` compares the latest version of a file against an
earlier vintage within the same release.

```{r}
# Compare latest vs. immediately previous vintage (default: version = -1)
vintage_changes <- compare_aux_vintages(
  measures = c("cpi", "gdp"),
  version  = -1
)

# Go two versions back
vintage_changes_v2 <- compare_aux_vintages(
  measures = "cpi",
  version  = -2,
  verbose  = TRUE
)
```

Each element of the result list contains:

- `diff_values`: data table of value-level differences (`NULL` if none).
- `diff_rows`: data table of added/removed rows (`NULL` if none).
- `key_cols`: character vector of primary key columns used for comparison.

If no previous version exists, all elements will be `NULL`.

---

# 5️⃣ Best Practices

- Always run updates via `update_aux_measures()` — it handles dependency
  order, status checks, and logging automatically.
- Use `log_save = TRUE` when running updates for a formal release so that
  the log is persisted to disk and retrievable later.
- Use log inspection (`aux_log_last()`, `pipfun::log_get()`) to verify
  that all measures updated successfully before proceeding.
- Skip `simulate_changes()` for actual production releases.
- Run `compare_aux_releases()` and `compare_aux_vintages()` after updates
  to confirm expected changes and detect unintended modifications.
- `compare_aux_releases()` requires a prior release with the same identity
  to exist on the Y-drive. If none is found, it will error.
