---
title: "Aux data -Pipeline Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{auxdata-pipeline-overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all()
#library(pipaux)
```

### Overview

-   **Objectives**: Update Auxiliary Data Files and Compare both across different releases and within a release

-   **Key functions**:

    -   `update_all_aux()`: update all auxiliary data automatically
    -   `compare_aux_releases()`
    -   `compare_vintage_versions()`

-   **Key Outputs**:

    -   Updated branches on GitHub and all auxiliary data files saved to the Y drive.\
        Note that the current folder used to store files from the new pipeline is defined by the option `getOption("pipaux.working_dir")`, which currently points to `"Y:/PIP_ingestion_pipeline_v2"`.

    -   Data tables showing changes between files in current release or bw diff releases

-   **Key Steps**:

    1.  Set up the working release
    2.  Update auxiliary data: in github and the Y drive
    3.  Compare auxiliary data files to identify changes

## STEP 1 - SETUP WORKING RELEASE

in the new pipeline framework, we always work with a reference version called the "working release." Therefore, we need to set up the working release at the beginning of each new R session:

In this example, I am using a "TEST" release with a release date of "20250203". However, if no release has been created for the desired date, call `pipfun::new_pip_release` with the appropriate arguments to create it.

``` r
# Set up the working release 
pipfun::setup_working_release(release  = "20250203",
                              identity = "TEST")
```

## STEP 2 - UPDATE ALL (OR SOME) AUXILIARY DATA

**Update**

``` r
# Update auxiliary data

# For the scope of this example I will rely on aux data repos in my GH account , thus I set owner "RossanaTat"

update <- update_all_aux(measures = c("maddison",
                                      "censoring",
                                      "pop"),
                         owner = "RossanaTat",
                         log = TRUE,
                         log_save = TRUE)
```

**Inspect log**

``` r
# Read log directly

pipfun::log_get(name = "pipaux_update_log")[]

# Interact with the log, e.g., filter certain events

pipfun::log_get(name = "pipaux_update_log")[event =="success"]

# Or you can read the log file saved in the folder
```

## Notes (for your reference or future modifications):

-   At this stage, I am working with auxiliary data repositories under my GitHub account. Once testing is complete, we will run the functions on the actual folders under the **"PIP-Technical-Team"** organization.

-   Additional functions to facilitate log interaction are under development. For now, we use those already available in the `{pipfun}` package.

    -   See the Log Vignette (here)[<https://pip-technical-team.github.io/pipfun/articles/log.html>]

## STEP 3 - COMPARE AUXILIARY DATA FILES (ACROSS RELEASES AND VERSIONS)

In this section, we explore how to identify changes in auxiliary data files. We focus on the following measures for this example: **cpi**, **pop**, **ppp**, **gdp**, and **pfw**.\

The current release used here is `"20250101_TEST"`.

**Compare between releases**

If `old_release = NULL`, the function will automatically detect the previous release.\

Alternatively, you can provide a specific release name, e.g., `"20250203_TEST"`.

``` r

# Identify changes in gdp

changes_gdp <- compare_aux_releases(measure = "gdp",
                                    old_release = NULL)

names(changes_gdp)

```

**Compare within releases**

By default, this compares the current file in the `aux_` folder with the previous version.\

Check the `version` argument documentation for more details.

``` r
changes_version_gdp <- compare_vintage_versions(measure = "gdp", version = -1)
```

## Notes (for your reference or future modifications):

These functions compare auxiliary data files either between releases or within the same release (i.e., across versions).

They are intended for use with files structured and saved through the new pipeline framework.

Since we haven’t run the full pipeline yet, I simulated changes (e.g., modified values, new rows or columns) in older files using helper functions for testing purposes.

As a result, the output values shown here are not real and stem from the simulated changes.


