---
title: "Aux Data - Pipeline Overview"
author: "PIP Technical Team"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aux Data - Pipeline Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
library(pipaux)
```

# Overview

This vignette provides an external-user view of the auxiliary data (aux data) pipeline, detailing how measures are updated, logged, and compared across releases and versions. It is intended for those who want to run updates safely, understand dependencies, and inspect changes over time.

## Objectives

- Update auxiliary data measures in dependency-aware order.
- Synchronize data between GitHub and Y-drive.
- Inspect logs for timing, success, and errors information.
- (Optionally simulate changes for testing purposes.)
- Compare data across releases or versions (vintages).


# Key Functions

| Function                | Purpose                                                                                  |
|-------------------------|------------------------------------------------------------------------------------------|
| `update_aux_measures()` | Updates all or selected auxiliary data measures in dependency order. Handles GitHub/Y-drive updates, logging, and dependency resolution. |
| `aux_fun()`             | Internal function called by `update_aux_measures()`. Executes measure updates and manages dependencies. |
| `check_status()`        | Determines whether updates are needed for a measure on GitHub or Y-drive.                || `get_fs_status()`       | Wrapper around `check_status()` to return update flags programmatically.           |
| `compare_aux_releases()`| Compares files across releases to detect new rows, removed rows, or changed values.      |
| `compare_aux_vintages()`| Compares files within the same release to detect version changes.                        |

# Key Outputs

- Updated GitHub release branches (`release_branch = <release>_<identity>`) for each measure.
- Synchronized auxiliary data files saved to Y-drive (`getOption("pipaux.working_dir")`).
- Logs capturing:
    - Measure name
    - Step type (update, check, error)
    - Timing and status
    - Optional error messages

# 1️⃣ Step 1 – Set up the Working Release

All operations in the pipeline are tied to a working release, which defines the release name and identity (e.g., TEST). Initialize at the start of each session:

```{r}
pipfun::setup_working_release(
  release  = "20260202", #for example
  identity = "TEST"
)
```

# 2️⃣ Step 2 – Update Auxiliary Data Measures

You can update all measures or a subset, in dependency-aware order, using `update_aux_measures()`. Logs are automatically captured.

```{r}
# Update specific measures with logging enabled
update_aux_measures(
  measures = c("cpi", "metaregion"),
  owner    = "RossanaTat",
  log      = TRUE,
  log_save = FALSE,        # Optional: save log to disk
  halt_on_dep_fail = FALSE # Continue updates even if a dependency fails
)
```

## Notes on Updates

- **Dependency handling:** Measures are updated in the correct order based on their dependencies.
- **GitHub/Y-drive updates:**  
    - `update_gh` flag controls GitHub branch updates.  
    - `update_y` flag controls saving updated data to the Y-drive.
- **Logging:**  
    - Use and inspect `aux_log_last()`, which provides quick access to the most recent log. Or read log saved to disk with `pipfun::log_load()`
- **Optional simulations:** For testing, you can simulate row/value changes, but this is not required for production updates.

```{r}
# Inspect log events
pipfun::log_get("pipaux_update_log")[event == "update"]
pipfun::log_get("pipaux_update_log")[event == "error"]
```


# 3️⃣ Step 3 – Simulate Changes (For TESTING ONLY)

Simulations allow you to test pipeline behavior by creating hypothetical row/value changes in auxiliary files:

```{r}
# Simulate changes for testing
simulate_changes(
  measures = c("cpi", "metaregion"),
  owner    = "RossanaTat",
  drop = c(1, 2),
  indices = c(1, 3)
)
```

**Important:** Simulations are optional and primarily for testing pipeline logic. For actual releases, ignore this step.

# 4️⃣ Step 4 – Compare Auxiliary Data

Once updates are applied, you can inspect changes across releases or within vintages.

## Compare Across Releases

```{r}
changes_pop <- compare_aux_releases(
  measure     = "pop"
)
```

## Compare Within the Same Release (Vintage Versions)

```{r}
vintage_changes_cpi <- compare_aux_vintages(
  measures = "cpi",
  version  = -1  # Compare to previous version
)

```

Each comparison provides:

- `diff_values`: Changed values
- `diff_rows`: Added/removed rows
- `key_cols`: Key cols used for comparison

If no previous version exists, results will be `NULL`.

# 5️⃣ Best Practices

- Always run updates in dependency order, relying on `update_aux_measures()`.
- Use logs to trace updates, timing, and errors. Use `{pipfun}` logging functions to inspect and save logs as needed, e.g., `pipfun::log_load()` and `pipfun::log_save()`, which are powered by `{stamp}`.
- Skip simulations for actual releases.
- Verify release/vintage comparisons to ensure data integrity.
- Optionally use `get_fs_status()` to programmatically verify if updates are needed for a specific measure. 
