---
title: "dependencies_workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dependencies_workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pipaux)
```

In this vignette, we explore how to update any of the auxiliary data measures. The process is specifically designed to automatically manage all interdependencies among auxiliary data, so that when updating a specific measure—such as `ppp`—all of its dependencies—such as `country_list`—are automatically updated first, if needed.

The main function you need to use to implement the full workflow is `aux_fun`. This is a general function that works with any of the auxiliary data measures and takes care of the whole update process. The process is based on 3 key steps:
1. checking if the measure has any dependency
2. Checking if any of the depndencies needs to be updated in GitHub and/or the Y: drive (MEMO: add details of what this means )
3. Update any dependency found in 2, in GitHub and/or Y: Drive  
4. Update the original measure 

## Preliminary operations

Remember that in the new pipeline framework, we always work with a reference version called the "working release." Therefore, before implementing the dependency workflow, it is important to set up the working release at the beginning of each new R session:

In this example, I am using a "TEST" release to demonstrate the new {pipaux} functionalities for managing dependencies with a test release of "20250203"

```{r}

# If no release has been created for the desired date, call pipfun::new_pip_release with the appropriate arguments to create it 
# new_release <- pipfun::new_pip_release()

# Set up the working release 
pipfun::setup_working_release(release  = "20250203",
                              identity = "TEST")

# Get the working release 
pipfun::get_wrk_release()
```

## Running an example

