---
title: "Auxiliary Data - Version Control"
subtitle: "Release branch management"
author: Rossana Tatulli
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auxiliary Data - Version Control}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(pipaux)
#devtools::load_all(".")
library(pipaux)
```

The {pipaux} package manages auxiliary data used across PIP. It includes functions like `aux_ppp()` for PPP data and `aux_gdp()` for GDP data (formerly pip_*()), which allow users to update or load specific auxiliary data. For example, calling `aux_gdp()` will update or load GDP data depending on user specification. When updating data, these functions push the changes to the DEV branch of the respective auxiliary data repository. The DEV branch of each auxiliary data repository is therefore updated with changes. 

To ensure that the version of auxiliary data used for each release is properly tracked, 2 steps are necessary:

- Each auxiliary data repository should have a dedicated **release branch**.
- The release branch must be periodically **updated** with changes from the **DEV branch**.
To facilitate the management of release branches and synchronization with the DEV branch, the {pipfun} package provides functions designed for this purpose.

Specifically, 

### 1. Check if release branch is there, and if not, create it 

```{r initial-setup}
# install.packages("pak")
#pak::pak("PIP-Technical-Team/pipfun@DEV_v2")

library("pipfun")

```


```{r}

# Get branches in repo

# Example 1: TEST repo
repo_branches <- pipfun::get_repo_branches(repo  = "aux_test", 
                                           owner = getOption("pipfun.ghowner"))
repo_branches

# Example 2: PPP repo

pipfun::get_repo_branches(repo  = "aux_ppp", 
                          owner = getOption("pipfun.ghowner"))


# As no release branch is found, create it from DEV
# create_new_branch(new_branch = "20241005",
#                   ref_branch = "DEV",
#                   repo = "aux_ppp")

```


### 2. Check that release branch is updated with DEV
```{r}

pipfun::compare_branch_content(repo    = "aux_test",
                               branch1 = "DEV",
                               branch2 = "20241005")

```

### 3. Update release branch if needed

In this case, given that the release branch has just been created, it is up to date with DEV. However, had "same_content" been `FALSE`, we need to update release branch accordingly as follows:


```{r}

# Option 1: Make release branch point to the latest commit as in DEV
pipfun::update_branches(repo    = "aux_test",
                        branch1 = "DEV",
                        branch2 = "20241005")

# Option 2: Merge DEV into release branch

pipfun::merge_branch_into(repo          = "aux_test",
                          source_branch = "DEV",
                          target_branch = "20241005")
```







