---
title: "Utilizing the merge function"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

library(pipaux)

```

# How to use the `merger_aux`

To efficiently merge auxiliary files, you can utilize the merger_aux() from pipaux package. This function requires three arguments: the two data files you want to merge and the type of merge to perform. The merge type has six options: left, right, full, using, master, and inner, with "left" being the default.

Here’s a breakdown of how it works:

1.  Arguments:

-   Data Files: You need to provide two auxiliary datasets that you want to merge. These are the primary inputs for the function.
-   Merge Type: This argument specifies the nature of the merge operation. You can choose from six different merge types:
    -   **Left**: Includes all records from the left dataset and matches records from the right dataset where possible.

    -   **Right**: Includes all records from the right dataset and matches records from the left dataset where possible.

    -   **Full**: Includes all records from both datasets, with matching records where available and missing records filled with nulls.

    -   **Using**: Merges the datasets based on specified columns or keys.

    -   **Master**: Uses the left dataset as the primary dataset and adds records from the right dataset.

    -   **Inner**: Includes only the records that are present in both datasets.

2.  Default Option: If you don’t specify a merge type, the function defaults to "left", meaning it will keep all records from the left dataset and include matching records from the right dataset where available.

**Note:** Keys/IDs and names of auxiliary data files (such as *cpi*, *ppp*, *pfw*, *pop*, *gdp*, *gdm*, etc.) are embedded within the data files as attributes. The `merger_aux()` function utilizes these IDs to create suitable key variables for merging the data files and generating an ID for the combined data file. Additionally, the function produces a new auxiliary file name by combining the names of the files being merged.

# Examples

### 1. Merge pfw and cpi files

```{r}

# load pfw data from the testing folder
pfw <- pip_pfw(action = "load",
               branch = "DEV",
               maindir = "//w1wbgencifs01/pip/PIP-Data_Testing/pipaux_Testing")

# load cpi
cpi <- pip_cpi(action = "load",
               branch = "DEV",
               maindir = "//w1wbgencifs01/pip/PIP-Data_Testing/pipaux_Testing")

cpi <- cpi[, -c("cpi_domain")]

pfw_cpi <- merger_aux(pfw, cpi)

```

```{r}

# load ppp wide format data
ppp <- pip_ppp(action = "load",
               branch = "DEV",
               maindir = "//w1wbgencifs01/pip/PIP-Data_Testing/pipaux_Testing")

pfw_cpi_ppp <- merger_aux(pfw_cpi, ppp)

```

